A simple and efficient algorithm for computing approximate Nash equilibria in sequential games with imcomplete information

### 2015-07-28 paper.tex
polyhedron changed to polytope (since bounded provably)

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e}
\usepackage{url}
\usepackage{sans}
\usepackage{amsmath,amsfonts, amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage[FIGTOPCAP]{subfigure}

@@

\subsection{Nash-equilibrium concepts and statement of the problem}
The sequence-form representation for two-person zero-sum games with
incomplete information was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium} where it was
established that for such games, there exist sparse matrices
$A \in \mathbb{R}^{n_1 \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times
  n_1}$, $E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in
\mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$,
$l_1$, and $l_2$ are all linear in the size of the game tree (number
of states in the game) and such that Nash equilibria correspond to
pairs $(x, y)$ of \textit{realization plans} which solve the primal
LCP (Linear Convex Program)
\begin{equation}
  \begin{aligned}
     \text{minimize }\langle e_1, p\rangle, \hspace{.5em}\text{
       subject to: }(y,p) \in \mathbb{R}^{n_2} \times
     \mathbb{R}^{l_1},  y \ge 0, E_2y = e_2, -Ay + E_1^Tp \geq 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

and the dual LCP
\begin{equation}
  \begin{aligned}
    \text{maximize}-\langle e_2, q\rangle,\hspace{.5em}\text{subject
      to: }(x,q) \in \mathbb{R}^{n_1} \times \mathbb{R}^{l_2},  x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, and each $e_k$ is a vector of the form
$(1, 0, ..., 0)$. In this so-called \textit{sequence-form}
representation, the strategy profile of player $k$ is the convex polyhedron
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k}_+ |\text{ }E_kz = e_k\}.
\label{eq:polyhedron}
\end{equation}

Note that the LCPs above have the equivalent saddle point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }\langle x, Ay\rangle
  \label{eq:gilpin}
\end{equation}

%% Note that $\langle e_1, p\rangle = p_0$ and $\langle e_2, q\rangle
%% = q_0$ by definition of $e_1$ and
%% $e_2$. 
At a feasible point $(y, p, x, q)$ for the LCPs, the \textit{primal-dual gap}
$\tilde{G}(y, p, x, q)$ of these primal-dual pair is given
by\footnote{The inequality being due to \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  0 \le \tilde{G}(y, p, x, q) &:= \langle e_1, p\rangle - (-\langle
  e_2, q\rangle) = \langle e_1, p\rangle + \langle
  e_2, q\rangle\\
  &= G(x, y) := \mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle |
(u,v) \in Q_1 \times Q_2\}.
\end{split}
  \label{eq:dgap}
\end{eqnarray}
In \eqref{eq:dgap}, the quantity $G(x, y)$ is nothing but the primal-dual
gap for the equivalent saddle point problem \eqref{eq:gilpin}.
It was shown (see Theorem 3.14 of \cite{vonequilibrium}) that a pair
$(x, y) \in Q_1 \times Q_2$ of realization plans is a solution to the
LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} (i.e is a Nash
equilibrium for the game)  if and only if there exist vectors $p$ and
$q$ such that
\begin{equation}
\hspace{.25em} -Ay + E_1^Tp \ge 0, \hspace{.5em}A^Tx + E_2^Tq \ge
0, \hspace{.25em} \langle x, -Ay + E_1^Tp\rangle = 0, \hspace{.25em}
\langle y, A^Tx  + E_2^Tq\rangle = 0.
\label{eq:feasibility}
\end{equation}

@@

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polyhedron} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polyhedral strategy profiles $Q_k$
defined in \eqref{eq:polyhedron} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Our contribution}
One cannot directly attack the LCPs \eqref{eq:primal_pb} and
\eqref{eq:dual_pb} via a traditional primal-dual algorithm
(for example \cite{chambolle2010,chambolle2014ergodic}) because
computing the euclidean projections $\proj_{Q_k}$ is very difficult
(in fact, such subproblems would have to be solved
iteratively\footnote{An ``exception to the rule'' is the case where
  the $Q_k's$ are simplexes, so that the projections can be computed
  exactly using \cite{duchi2008efficient}.}). Also,
the primal-dual gap might explode even at points arbitrarily close to
the set of feasible points, leaving the algorithm with no indication
whatsoever, on whether progress is being made or not. So need to way to
\begin{itemize}
\item[--] avoid having to compute the projections $\proj_{Q_k}$,
\item[--] have control over how far we are from the set of equilibria and
  avoid infinite (and thus non-informative) primal-dual gaps.
\end{itemize}

@@

The primal-dual algorithm first developed in \cite{chambolle2010}, was
proposed can be used to solve matrix games on
simplexes as was done in \cite{chambolle2014ergodic}, for example.
It should be stressed that such matrix games are
considerably simpler than the games considered here. Indeed, the
authors in \cite{chambolle2014ergodic} use the fact that computing the
euclidean projection of a point onto a simplex can be done in linear
time as in \cite{duchi2008efficient}. In contrast, no such efficient
algorithm is known nor is likely to exist for the polyhedra $Q_k$
defined in \eqref{eq:polyhedron}, the strategy profiles for players in
the games considered here. It should however be noted that such this
projection can still be done iteratively using, for example, the
algorithm in proposition 4.2 of \cite{combettes2010dualization} or the
algorithms developed in \cite{tran2015splitting}. Unfortunately, as
with any nested iterative scheme, one would have to solve this
sub-problem with finer and finer precision.

@@

By Theorem \ref{thm:pd}, solviing for a Nash equilibrium for
the game is equivalent to solving the GSP
\eqref{eq:unconstrained_pb}, which as it turns out, is simpler
conceptually. The rest of the paper will be devoted to
developing an algorithm for solving the latter.

@@

\begin{theorem}[Ergodic $\mathcal{O}(1/\epsilon)$ convergence]
Let $d_0$ be the euclidean distance between the starting point
$(y^{(0)},p^{(0)},x^{(0)},q^{(0)})$ of Algorithm \ref{Tab:algo} and the
set of equilibria for the GSP \eqref{eq:unconstrained_pb}.
Then given any $\epsilon > 0$, there exists an index
$k_0 \le \frac{2d_0\|K\|}{\epsilon}$ such that after $k_0$ iterations
the algorithm produces the following a quadruplet
$(y^{k_0},p^{k_0},x^{k_0},q^{k_0})$ and a vector $v^{k_0}$ such that
\begin{eqnarray}
\|v^{k_0}\| \le \epsilon \text{ and}\hspace{.5em} v^{k_0} \in
\partial[\hat{\Psi}_1(., ., x^{k_0}, q^{k_0}) +
  \hat{\Psi}_2(y^{k_0}, p^{k_0}, ., .)](y^{k_0},p^{k_0},x^{k_0},q^{k_0}).
\end{eqnarray}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e}
\usepackage{url}
\usepackage{sans}
\usepackage{amsmath,amsfonts, amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage[FIGTOPCAP]{subfigure}
\usepackage{caption}

@@

\subsection{Nash-equilibrium concepts and statement of the problem}

The sequence-form representation for two-person zero-sum games with
incomplete information was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium} where it was
established that for such games, there exist sparse matrices
$A \in \mathbb{R}^{n_1 \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times
  n_1}$, $E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in
\mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$,
$l_1$, and $l_2$ are all linear in the size of the game tree (number
of states in the game) and such that Nash equilibria correspond to
pairs $(x, y)$ of \textit{realization plans} which solve the primal
LCP (Linear Convex Program)
\begin{equation}
  \begin{aligned}
     \text{minimize }\langle e_1, p\rangle, \hspace{.5em}\text{
       subject to: }(y,p) \in \mathbb{R}^{n_2} \times
     \mathbb{R}^{l_1},  y \ge 0, E_2y = e_2, -Ay + E_1^Tp \geq 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

and the dual LCP
\begin{equation}
  \begin{aligned}
    \text{maximize}-\langle e_2, q\rangle,\hspace{.5em}\text{subject
      to: }(x,q) \in \mathbb{R}^{n_1} \times \mathbb{R}^{l_2},  x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, with exactly 1 entry equal -1 per row, and each $e_k$ is a vector of the form
$(1, 0, ..., 0)$.

\begin{figure}[!h]
\begin{minipage}{.65\linewidth}
%\caption*{\textbf{Sparse-Variation.}
Note that the LCPs above have the equivalent saddle point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }\langle x, Ay\rangle,
  \label{eq:gilpin}
\end{equation}
where the \textit{convex polytope} (by which we mean a bounded convex
polyhedron)
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k}_+ |\text{ }E_kz = e_k\} \subseteq
  [0, 1]^{n_k}
\label{eq:polytope}
\end{equation}
is the the strategy profile of player $k$ in this representation.
At a feasible point $(y, p, x, q)$ for the LCPs, the \textit{primal-dual gap}
$\tilde{G}(y, p, x, q)$ of these primal-dual pair is given
by\footnote{The inequality being due to \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  0 &\le \tilde{G}(y, p, x, q) := \langle e_1, p\rangle - (-\langle
  e_2, q\rangle) = \langle e_1, p\rangle + \langle
  e_2, q\rangle\\
  &= G(x, y) := \mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle |
(u,v) \in Q_1 \times Q_2\}.
\end{split}
  \label{eq:dgap}
\end{eqnarray}
%}
\end{minipage}
\hspace{.5em}
\begin{minipage}{.4\linewidth}
  \includegraphics[width=1\linewidth]{grain4.png}
  \caption{A convex polytope}
  \label{fig:roi}
\end{minipage}
\end{figure}

In \eqref{eq:dgap}, the quantity $G(x, y)$ is nothing but the primal-dual
gap for the equivalent saddle point problem \eqref{eq:gilpin}.
It was shown (see Theorem 3.14 of \cite{vonequilibrium}) that a pair
$(x, y) \in Q_1 \times Q_2$ of realization plans is a solution to the
LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} (i.e is a Nash
equilibrium for the game)  if and only if there exist vectors $p$ and
$q$ such that
\begin{equation}
\hspace{.25em} -Ay + E_1^Tp \ge 0, \hspace{.5em}A^Tx + E_2^Tq \ge
0, \hspace{.25em} \langle x, -Ay + E_1^Tp\rangle = 0, \hspace{.25em}
\langle y, A^Tx  + E_2^Tq\rangle = 0.
\label{eq:feasibility}
\end{equation}

@@

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polytope} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polyhedral strategy profiles $Q_k$
defined in \eqref{eq:polytope} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Our contribution}
One cannot directly attack the LCPs \eqref{eq:primal_pb} and
\eqref{eq:dual_pb} via a traditional primal-dual algorithm
(for example \cite{chambolle2010,chambolle2014ergodic}) because
computing the orthogonal projections $\proj_{Q_k}$ is very difficult
(in fact, such subproblems would have to be solved
iteratively\footnote{An ``exception to the rule'' is the case where
  the $Q_k's$ are simplexes, so that the projections can be computed
  exactly using \cite{duchi2008efficient}.}). Also,
the primal-dual gap might explode even at points arbitrarily close to
the set of feasible points, leaving the algorithm with no indication
whatsoever, on whether progress is being made or not. So need to way to
\begin{itemize}
\item[--] avoid having to compute the projections $\proj_{Q_k}$,
\item[--] have control over how far we are from the set of equilibria and
  avoid infinite (and thus non-informative) primal-dual gaps.
\end{itemize}

@@

The primal-dual algorithm first developed in \cite{chambolle2010}, was
proposed can be used to solve matrix games on
simplexes as was done in \cite{chambolle2014ergodic}, for example.
It should be stressed that such matrix games are
considerably simpler than the games considered here. Indeed, the
authors in \cite{chambolle2014ergodic} use the fact that computing the
orthogonal projection of a point onto a simplex can be done in linear
time as in \cite{duchi2008efficient}. In contrast, no such efficient
algorithm is known nor is likely to exist for the polyhedra $Q_k$
defined in \eqref{eq:polytope}, the strategy profiles for players in
the games considered here. It should however be noted that such this
projection can still be done iteratively using, for example, the
algorithm in proposition 4.2 of \cite{combettes2010dualization} or the
algorithms developed in \cite{tran2015splitting}. Unfortunately, as
with any nested iterative scheme, one would have to solve this
sub-problem with finer and finer precision.

@@

By Theorem \ref{thm:pd}, solviing for a Nash equilibrium for
the game is equivalent to solving the GSP
\eqref{eq:unconstrained_pb}, which as it turns out, is simpler
conceptually (for example, we no longer need to compute the
complicated orthogonal projections $\proj_{Q_k}$). The rest of the
paper will be devoted to developing an algorithm for solving the latter.

@@

\begin{theorem}[Ergodic $\mathcal{O}(1/\epsilon)$ convergence]
Let $d_0$ be the orthogonal distance between the starting point
$(y^{(0)},p^{(0)},x^{(0)},q^{(0)})$ of Algorithm \ref{Tab:algo} and the
set of equilibria for the GSP \eqref{eq:unconstrained_pb}.
Then given any $\epsilon > 0$, there exists an index
$k_0 \le \frac{2d_0\|K\|}{\epsilon}$ such that after $k_0$ iterations
the algorithm produces the following a quadruplet
$(y^{k_0},p^{k_0},x^{k_0},q^{k_0})$ and a vector $v^{k_0}$ such that
\begin{eqnarray}
\|v^{k_0}\| \le \epsilon \text{ and}\hspace{.5em} v^{k_0} \in
\partial[\hat{\Psi}_1(., ., x^{k_0}, q^{k_0}) +
  \hat{\Psi}_2(y^{k_0}, p^{k_0}, ., .)](y^{k_0},p^{k_0},x^{k_0},q^{k_0}).
\end{eqnarray}
